---
title: "Final Project Preliminary EDA"
name: "Sarah Petro"
date: "November 15 2020"
output: html_notebook

---

##Research Question: 
What video games were the most popular in 2019? This will be based on sale numbers and I also plan to break the popularity down by genre and sales in Europe and the US. Finally, I will use critic and user scores to see which popular games by sales were rated the highest and lowest.

I found the data sets on Kaggle and downloaded the .csv files from the Kaggle website.


##Video Game Sales 2019
The first data set I used is Video Game Sales 2019 from Kaggle. It was gathered by Abdulshaheed Alqunber who is a student at Boston University and can be found at: https://www.kaggle.com/ashaheedq/video-games-sales-2019. He used this data set to continue the collection of data from the second dataset I used, Video Game Sales. Since the datasets are very similar, I plan to use the added benefit of looking specifically at 2019 data. However, this dataset also includes critic and user scores which I plan to use in my exploration.
Below, you can see a snippet of the data along with some summary statistics. 
```{r}
head(vgsales.12.4.2019)
```
```{r}
summary(vgsales.12.4.2019)
```
This dataset was originally collected as a continuation work of the second dataset I used, Video Game Sales. He explains that he scraped the data using a script and collected it into a .csv file.

A case in this dataset describes a video game and its respective sales data that includes the year 2019, which the first dataset does not include. Each video game is its own separate case.

I plan to use the Name, Genre, Critic_Score, and User_Score from this dataset to explore the most popular titles of 2019.

```{r}
nrow(vgsales.12.4.2019)
```
The dataset has a lot of rows for exploration at 55,792 rows in total.


##Video Game Sales
The second dataset I 